Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-03-23 src/ms.tex
referee comments

\subsection{Comparison with asteroseismic rotation rates}
\label{subsec:asteroseismic}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
    \caption{Comparison of CKS (left) and \edit1{\lamostmcq} samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane. Note, the CKS, LAMOST, and \citet{Hall2021} samples derive \teff from distinct pipelines. A constant offset of +41~K was applied to the LAMOST \teff to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}). Using LAMOST \teff, where it exists, for the \citet{Hall2021} sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
    \label{fig:asteroseismic}
\end{figure}

@@

We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages). However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties. The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}). In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun (using CKS ages), and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.

@@

\edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge. To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold. Furthermore, two independent samples with different observational biases than the rotation period catalogs considered here also appear to yield a pile-up of stars at the long-period edge. These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii.}

@@

\begin{figure}
    \includegraphics[width=\linewidth]{./figures/gaia.pdf}
    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = 0.5 \approx \mathrm{Ro}_\odot$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$. The top rows show 2–d histograms of the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data. Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above. The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.}
    \label{fig:gaia}
\end{figure}

>>>>

\subsection{Comparison with \edit1{literature surveys}}
\label{subsec:asteroseismic}

@@

\edit1{\citet{Masuda2021} inferred the \prot distribution of 144 late-F/early-G dwarfs in the Kepler field from their precise stellar radii and spectroscopically determined \vsini. As with the asteroseismic sample, we compare the CKS and \lamostmcq samples with the results of those authors. We find that the long-period pile-up we observe appears to overlap with the \citet{Masuda2021} data. We therefore consider it likely that the long-period pile-up is the same feature, or an extension of the same feature, that was detected in the \citet{Hall2021} and \citet{Masuda2021} studies.}

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{./figures/asteroseismic.pdf}
    \caption{Comparison of CKS (left) and \edit1{\lamostmcq} samples (right) with the \citet{Hall2021} main-sequence asteroseismic sample in the \teff--\prot plane. Note, the CKS, LAMOST, and \citet{Hall2021} samples derive \teff from distinct pipelines. A constant offset of +41~K was applied to the LAMOST \teff to bring that sample onto the same scale as the asteroseismic sample (see Appendix~\ref{app:teff}). Using LAMOST \teff, where it exists, for the \citet{Hall2021} sample brings that sample into even closer agreement with the long-period pile-up in the LAMOST-McQuillan sample.}
    \label{fig:asteroseismic}
\end{figure*}

@@

We determined that 90\% of the stars on the ridge have ages between 1.4--5.6~Gyr (using ages from the CKS catalog), or 2.3--5.9~Gyr (using SPOCS ages). \edit1{These ranges are consistent with the range of ages observed in the \citet{Hall2021} and \citet{Masuda2021} samples, as shown in Figure~11 of the latter reference.} However, we note that systematic effects in surveys and theoretical models lead to large uncertainties in isochrone ages that are not necessarily represented by the reported age uncertainties. The SPOCS ages in particular are questionable given that only $\sim$1\% of the stars have ages $<$~2~Gyr and that stars are concentrated at the upper age boundary for a given \teff (Figure~\ref{fig:ages}). In fact, there are only a handful of long-period pile-up stars in the CKS sample with ages older than the age of the Sun (using CKS ages), and almost all of the pile-up stars would be compatible with an age equal to or less than the Sun's given the large age uncertainties.

@@

\edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge. To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold. Furthermore, two independent samples with different observational biases than the rotation period catalogs considered here also appear to yield a pile-up of stars at the long-period edge. These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii. We do note, however, that we can not rule out the possibility that the long-period pile-up is wider in period space than we observe, with stars just above the pile-up being undetected through rotational brightness modulations.}

@@

\begin{figure}
    \includegraphics[width=\linewidth]{./figures/gaia-mcquillan.pdf}
    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = 0.5 \approx \mathrm{Ro}_\odot$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$. The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data. Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above. The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.}
    \label{fig:gaia-mcquillan}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{./figures/gaia-santos.pdf}
    \caption{The same as Figure~\ref{fig:gaia-mcquillan} but for the \citet{Santos2021} sample.}
    \label{fig:gaia-santos}
\end{figure}
